### 2023-02-21 manuscript/selbias_template.Rmd

### Pub bias {{i}}

According to the selection model, there was `r c("no", "significant")[(mlm[["{{i}}"]][["pubbias_sel"]][["LRTp"]] < .05)+1]` evidence of publication bias, LRT(2) `r report(mlm[["{{i}}"]][["pubbias_sel"]][["LRT"]])`, p `r report(mlm[["{{i}}"]][["pubbias_sel"]][["LRTp"]])`.

Egger's test for funnel plot asymmetry showed `r c("no", "significant")[(mlm[["{{i}}"]][["pubbias"]][["pval"]] < .05)+1]` evidence of publication bias, Z `r report(mlm[["{{i}}"]][["pubbias"]][["zval"]])`, p `r report(mlm[["{{i}}"]][["pubbias"]][["pval"]])`.

```{r pubbbias{{i}}}
knitr::include_graphics("funnel_{{i}}.png")
```